Rectangulotopes

### 2024-04-03 draft/rectangulotopes.tex

In order to exhibit a similar realization for the strong rectangulotopes, we need to count subsets of leaves that lie in some subtrees of both the source and target trees of a rectangulation.
Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
We denote by %$i|j$ and $\frac{\ j\ }i$
$i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.

@@

\begin{proof}
  From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}. For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have
  \[
  \SP(A) = \{ x\in\R^n : \sum_{i\in S} x_i \leq \loday{f}_I(S) \text { and } \sum_{i\in [n]} x_i = \loday{f}_I([n]) \},
  \]
  where
  \[
  \loday{f}_I(S) := \llbracket S\cap I\not=\emptyset \rrbracket .
  \]
  Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
  \[
  \antiloday{f}_I(S) := - \llbracket S\subseteq I \rrbracket .
  \]
  Their sum
  \[
  f(S) := \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
  \]
  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
\end{proof}

>>>>

In order to exhibit a similar realization for the strong rectangulotopes, we need to count subsets of leaves that lie in some subtrees of both the source and target trees of a rectangulation.
Two leaves of the trees $T(R)$ and $S(R)$ are said to be \defn{common leaves} if the edges to their parents lie on the same segment of $R$.
We denote by %$i|j$ and $\frac{\ j\ }i$
$i \, \backslash \, j$ the situation where the rectangle labeled $i$ touches the rectangle labeled $j$ either from the left or from below in $R$. %, respectively.
We also use the Iverson bracket $\llbracket \varphi\rrbracket$, which is equal to 1 if $\varphi$ holds, and 0 otherwise.

@@

\begin{proof}
  From Lemmas~\ref{lem:weakMinkowski} and~\ref{lem:submodsum}, the submodular function defining the realization of $\WRP(n)$ is the sum of the submodular functions defining the Loday and anti-Loday shard polytope in Lemmas~\ref{lem:lodaysp} and~\ref{lem:antilodaysp}. For a Loday arc $A$ defined by an interval $I$ of $[n]$, we have that its shard polytope $\SP(A)$ is realized by the submodular function
  \[
  \loday{f}_I(S) := \llbracket S\cap I\not=\emptyset \rrbracket .
  \]
  (Recall that $\llbracket \varphi\rrbracket$ is equal to 1 if $\varphi$ holds, and 0 otherwise.)
  Similarly, for an anti-Loday arc $A$ defined by an interval $I$, the shard polytope $\SP(A)$ is defined by the following submodular function:
  \[
  \antiloday{f}_I(S) := - \llbracket I\subseteq S \rrbracket .
  \]
  Their sum
  \[
  f(S) := \sum_{I\text{ interval of }[n]} \loday{f}_I(S) + \antiloday{f}_I(S)
  \]
  is therefore exactly the number of intervals of $I$ that intersect $S$ but are not contained in $S$, hence the number of intervals intersecting both $S$ and its complement.
\end{proof}
